Non-word repetition in Yélî Dnye

### 2020-12-19 manuscript.tex
minor typo fixes, pval replacement, & prettier phone list printing
also i decided on a version of option 2 for deletions/substitutions

We ran children in batches, testing within a handful of hamlet clusters
spread across the northeastern region of the island. Because space
availability was limited in different ways from hamlet to hamlet, the
places where elicitation happened varied across testing sites. We tested
in four different sites, only making a single visit to each, conducting
back-to-back testing of all eligible children present at the time of our
visit in order to prevent the items `spreading' between children through
hearsay. In the first hamlet, we tested children in five different
places, with some children being tested inside a house and others tested
on the veranda. The complete list of places and the ways in which they
met the desiderata mentioned above can be found in the raw data,
available from online supplementary materials. Whenever children living
in the same household were tested, we tried to test children in age
order, from oldest to youngest, to minimize intimidation for younger
household members, and always using different elicitaiton sets.

We fitted the child with a headset microphone (Shure SM10A or WH20 XLR
with a dynamic microphone on a headband, most children using the former)
that fed into the left channel of a Tascam DR40x digital audio recorder.
The headsets were designed for adult use and could not be comfortably
seated on many children's heads without a more involved adjustment
period. To minimize adjustment time, which was uncomfortable for some
children given the proximity of the experimenter and equipment, we
placed the headband on children's shoulders in these cases, carefully
adjusting the microphone's placement so that it was still close to the
child's mouth. A research assistant who spoke Yélî Dnye natively sat
next to the child throughout the task to provide instructions and, if
needed, encouragement. The research assistant coached the child
throughout to task to make sure that they understood what they were
expected to do. An experimenter (the first author) delivered the
pre-recorded stimuli to the research assistant and the child over
headphones.

The first phase of the experiment involved making sure the child
understood the task. We explained the task orally presented the first
practice item. At this point, many children did not say anything in
response, at which point we initiated the following procedure: First,
the assistant insisted the child make a response. If the child still did
not say anything, the assistant said a real word and then asked the
child to repeat it, then another and another. If the child could repeat
real words correctly, we provided the first training item over
headphones again for children to repeat. Most children successfully
started repeating the items at this point, but a few needed further
help. In this case, the assistant modeled the behavior (i.e., the child
and assistant would hear the item again, and the assistant would repeat
it; then we would play the item again and ask the child to repeate it).
A small minority of children still failed to repeat the item at this
point. If so, we tried again with the second training item, at which
point some children got it and could continue. A fraction of the
remaining children, however, failed to repeat this second training item,
as well as the third one, in which case we stopped testing altogether
(see Participants section for exclusions).

The second phase of the experiment involved running the child through
the list of test items randomly assigned to them. This was done in the
same manner as the practice items: the stimulus was played over the
headphones, and then the child repeated it aloud. NWR studies vary in
whether children are allowed to hear and/or repeat the item more than
one time. We had a fixed procedure for the test items (i.e., the
non-practice items) in which the child was alowed to make a further
attempts if their first attempt was judged erroneous in some way by the
assistant. The procedure worked as follows: When the child made an
attempted a repeat, the assistant indicated to the experimenter whether
the child's production was correct or not. If correct, the experimenter
would whisper this note of correct repetition into a separate headset
that fed into the right channel of the same Tascam and the group moved
on to the next item. If not, the child was allowed to try again, with up
to five attempts allowed before moving on to the next item. Children
were not asked to make repetitions if they did not produce a first
attempt.

In total, test sessions took approximately XX minutes on average, with
XX of those minutes attributed to practice and XX to the actual test
list.

@@

Additionally, we estimated the typological frequency of all phonological
segments used in the target items using the PHOIBLE cross-linguistic
phonological inventory database (Moran \& McCloy, 2019). For each phone
in our task, we extracted the number and percentage of languages noted
to have that phone in its inventory. While PHOIBLE is an unprecedentedly
comprehensive database, with phonological inventory data for over 2000
languages at the time these data were analyzed, it is of course still
far from complete, so estimated frequencies should be taken with a pinch
of salt. In both PHOIBLE and UPSID---another large phonemic inventory
database---note that nearly half of the segment types only appear in a
single language (Steven Moran, personal communication; see also
\href{http://web.phonetik.uni-frankfurt.de/upsid_info.html}{web.phonetik.uni-frankfurt.de/upsid\_info.html}).
It is therefore unsurprising that we did not find three of our segments
(/lβʲ/, /ṭp/, and /tp/) in the database. Following the general pattern
that many phones only appear once cross-linguistically, even in large
databases, we treat our ``missing'' phones as having a frequency of 1,
with a (rounded) percentage of 0\%.

@@

We tested a total of 55 children from 38 families spread across four
hamlets. We excluded test sessions from analysis for the following
reasons: refused participation or failed to repeat items presented over
headphones even after coaching (N=8), spoke too softly to allow offline
coding (N=5), or were 13 years old or older (N=2; we tested these
teenagers to put younger children at ease). The remaining 40 children
(14 girls) were aged 6.98 years (range 3.92-11.03 years). Among these
children were 34 only exposed to Yélî Dnye at home and 6 children
exposed to one or more other languages at home. Maternal years of
education averaged 8.22 years (range 6-12 years).\footnote{Years of
  education is difficult to precisely pinpoint in this context.
  Schooling beyond grade 8 takes place off-island and enrollment is tied
  to an ability to pay that year's school fees, seasonal interruptions
  to school opening times, familial obligations, and more. We asked
  parents to report their highest completed level of education. We then
  record the number of years entailed by having completed that level
  under ideal conditions. This method results in many esimates around
  typical ``graduation'' times (e.g., graduating from primary education
  before starting secondary education).} In terms of birth order, 6 were
first borns, 5 second, 2 third, 7 forth, 5 fifth, and 1 sixth.

@@

We first checked whether whole-item NWR scores varied between first and
subsequent presentations of an item by averaging word-level scores at
the participant level separately for first attempts and subsequent
repetitions. We excluded 1 children who did not have data for one of
these two types. As shown in Figure \ref{fig:Fig1-first-vs-others},
participants' mean word-level scores became more heterogeneous in
subsequent repetitions. Surprisingly, whole-item NWR scores for
subsequent repetitions (M = 40, SD = 26) were on average lower than
first ones (M = 64, SD = 15), t(38) = 6.28, p = 0; Cohen's d = 1.14).
Given uncertainty in whether previous work used first or all
repetitions, and given that score here declined and became more
heterogeneous in subsequent repetitions, we focus the remainder of our
analyses only on first repetitions.

Taking into account only the first attempts, we averaged attempts by
each of the 40 children who had data for first attempts; their ages
ranged from 3 to 10 (M = 6.50, SD = 1.50).

The overall NWR score was M = 64\% (SD = 15\%), Cohen's d = 4.36. Scores
based on phonemes are even higher M = 78\% (SD = 9\%), Cohen's d = 8.55.
The phoneme-based normalized Levenshtein distance was M = 22\% (SD =
9\%), meaning that about a fifth of phonemes were were substituted,
inserted, or deleted (the normalized Levenshtein distance is the
complement of phoneme-based scores).

@@

In this analysis, we were interested in variation in whole-item NWR
scores as a function of the average frequency with which sounds
composing individual target words are found in languages over the world.
To look at this, we fit a mixed logistic regression, in which the
outcome variable was whether the non-word was correctly repeated or not.
The fixed effects of interest was the average cross-linguistic phone
frequency; we also included child age as a control fixed effect, and
child ID \& target ID as random variables.

We could include 850 observations, from 40 children producing in any
given trial one of 41 potential target words. The analysis revealed a
main effect of age (ß = 0.33, SE ß = 0.13, p = 0.01); and a significant
estimate for the scaled average cross-linguistic frequency of phones in
the target words (ß = 0.82, SE ß = 0.18, p = 0): Target words with
phones found more frequently across languages had a higher proportion of
words that were correctly repeated, as clear in Figure 2. Averaging
across items and participants, the Pearson correlation between scaled
average cross-linguistic phone frequency and whole-item NWR scores was r
r(31) = 0.63.

@@

We next checked whether the association between whole-item NWR scores
and cross-linguistic phone frequence is actually be due to frequency of
the sounds within the language: One can suppose that sounds that occur
more frequently across languages are also more frequent within a
language, and therefore may be easier for children to represent and
repeat. We estimated frequency of the phones present in the stimuli in a
corpus of child-centered recordings (Casillas et al., 2020) by counting
the number of word types in which they occurred, and applied the natural
logarithm.\footnote{We also carried out this analysis using token phone
  frequency, but this measure was not correlated with whole-item NWR
  scores, and therefore the fact that it did not explain away the
  predictive value of cross-linguistic phone frequency was less
  informative than the relationship discussed in the main text, with
  type frequencies.} As before, unattested sounds were not considered
(i.e., they were declared NA so that they do not count for analyses).
Phone frequencies estimated from this corpus were correlated with
cross-linguistic phone frequencies (r(27)=0.52, =0.00). Moreover,
averaging across items and participants, the Pearson correlation between
scaled average corpus phone frequency and whole-item NWR scores was r =
0.46. Therefore, we fit another mixed logistic regression, this time
declaring as fixed effects both scaled cross-linguistic and corpus
frequencies (averaged across all attested phones within each stimulus
item), in addition to age. As before, the model contained random slopes
for both child ID and target. In this model, both cross-linguistic
frequency (ß = 0.82, SE ß = 0.27, p = 0.00) and age (ß = 0.33, SE ß =
0.13, p = 0.01) were significant predictors of whole-item NWR scores,
but corpus frequency (ß = 0.00, SE ß = 0.25, p = 1.00) was not.

@@

\textbf{Option 1: just numbers, first attempts only } Next, we
investigated patterns of deletion and substitution. Deletions were
relatively rare, with only 7 vowels deleted, and 3 consonants.

As for substitutions, it was as common for a nasal vowel to be produced
as an oral vowel as vice versa (34 oral target vowels produced as nasal
vowels, 30 nasal target vowels produced as oral vowels). Substitutions
in which the oral nature was preserved but the quality of the vowel was
changed were a great deal more common than changes in quality among
nasal vowels (110 oral vowels produced with a different quality; 13
nasal vowels produced with a different quality). As for consonants,
asymmetries were very marked with more complex consonants (specifically
dptpkpkmknmbghlv) mispronounced as simple ones (specifically
mńlwyvdgptkfhch, 115) than vice versa (2). Simple consonants were
mispronounced as other simple consonants quite frequently (68 simple
consonants mispronounced as other simple ones, compared to 33 complex
ones).

\textbf{Option 2: numbers and proportions, first attempts only} Next, we
investigated patterns of deletion and substitution. Deletions were
relatively rare, with only 7 vowels deleted (about 0.23\% of all vowel
targets), and 3 consonants deleted (about 0.12\% of all consonant
targets).

As for substitutions, it was as common for a nasal vowel to be produced
as an oral vowel as vice versa (34 oral target vowels produced as nasal
vowels or about 0.99\% out of all oral vowel targets, 30 nasal target
vowels produced as oral vowels or about 2.37\% out of all nasal vowel
targets). Substitutions in which the oral nature was preserved but the
quality of the vowel was changed were a great deal more common than
changes in quality among nasal vowels (110 oral vowels produced with a
different quality or about 1.86\% out of all oral vowel targets; 13
nasal vowels produced with a different quality or about 2.06\% out of
all nasal vowel targets). As for consonants, asymmetries were very
marked with more complex consonants (specifically dptpkpkmknmbghlv)
mispronounced as simple ones (specifically mńlwyvdgptkfhch, 115 times or
about 1.96\% out of all complex consonant targets) than vice versa (2
times or about 1.96\% out of all simple consonant targets). Simple
consonants were mispronounced as other simple consonants quite
frequently (68 simple consonants mispronounced as other simple ones or
about 0.38\% out of all simple consonant targets, compared to 33 complex
ones or about 1.02\% out of all complex consonant targets).

\textbf{Option 3: numbers and proportions, ALL attempts} Next, we
investigated patterns of deletion and substitution. Deletions were
relatively rare, with only 12 vowels deleted (about 0.28\% of all vowel
targets), and 6 consonants deleted (about 0.19\% of all consonant
targets).

As for substitutions, it was as common for a nasal vowel to be produced
as an oral vowel as vice versa (52 oral target vowels produced as nasal
vowels or about 1.10\% out of all oral vowel targets, 58 nasal target
vowels produced as oral vowels or about 2.53\% out of all nasal vowel
targets). Substitutions in which the oral nature was preserved but the
quality of the vowel was changed were a great deal more common than
changes in quality among nasal vowels (205 oral vowels produced with a
different quality or about 2.16\% out of all oral vowel targets; 23
nasal vowels produced with a different quality or about 2.36\% out of
all nasal vowel targets). As for consonants, asymmetries were very
marked with more complex consonants (specifically dptpkpkmknmbghlv)
mispronounced as simple ones (specifically mńlwyvdgptkfhch, 250 times or
about 2.43\% out of all complex consonant targets) than vice versa (2
times about 2.43\% out of all simple consonant targets). Simple
consonants were mispronounced as other simple consonants quite
frequently (120 simple consonants mispronounced as other simple ones or
about 0.52\% out of all simple consonant targets, compared to 55 complex
ones or about 0.98\% out of all complex consonant targets).

Finally, we looked into the frequency with which mispronunciations
resulted in real words. Two thirds of them were: 63\%. This is a type of
analysis that is seldom reported, and we could only find one comparison
point: ({\textbf{???}}) found that illiterate adults NWR
mispronunciations resulted in real words in 11.16\% of cases, whereas
literate adults did so in only 1.71\%. The percentage we observe here is
much higher than this, but we do not know whether age, language, or even
test structure explains this difference.

@@

Next, we inspected whether NWR scores varied as a function of word
length. Results are shown on table XX. This table shows that scores for
monosyllables was much lower than for other lengths. This is likely due
to the fact that the majority of monosyllables included were chosen
because they had sounds that are rare in the world's languages, which
may indicate that they are hard to produce or to perceive.

Setting monosyllables aside, we observe the typical pattern of lower
scores for longer items, although this is particularly salient for the
whole word scoring. This is the most commonly reported type of score,
but it is also the least forgiving. The pattern is less marked when
other two scores are used, which are less sensitive to errors. Averaging
across participants and items, the Pearson correlation between length
(2-4 syllables) and whole-item NWR scores was r(1) = -0.91. In a
generalized binomial mixed model, we included 479 observations, from 40
children producing in any given trial one of 24 potential target words.
The analysis revealed a main effect of age (ß = 0.56, SE ß = 0.14, p =
0); and a significant estimate for the length of the target words in
number of syllables (ß = -0.15, SE ß = 0.33, p = 0.65).

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Fig2-scores by age-1.pdf}
\caption{(\#fig:Fig2-scores by age)NWR whole-word scores for individual
participants as a function of age and sex (blue = boys, pink = girls).}
\end{figure}

Our final exploratory analysis assessed whether variation in scores was
structured by factors that vary across individuals. As shown in Figure
2, there was a greater deal of variance in earlier than later ages, with
significantly higher NWR scores for older children (Spearman's rank
correlation, given inequality of variance, rho (6,014.70) = 0.44, p =
0.00). In contrast, there was no clear association between NWR scores,
on the one hand, and sex (t (-0.29) = 27.56, p = 0.77), birth order
(data missing for 15 children, rho = (3,441.90) = -0.18, p = 0.39) or
maternal education (data missing for 0 children, rho = ( 9,594.37) =
0.10, p = 0.54).

>>>>

We ran children in batches, testing within a handful of hamlet clusters
spread across the northeastern region of the island. Because space
availability was limited in different ways from hamlet to hamlet, the
places where elicitation happened varied across testing sites. We tested
in four different sites, only making a single visit to each, conducting
back-to-back testing of all eligible children present at the time of our
visit in order to prevent the items from `spreading' between children
through hearsay. In the first hamlet, we tested children in five
different places, with some children being tested inside a house and
others tested on the veranda. The complete list of places and the ways
in which they met the desiderata mentioned above can be found in the raw
data, available from the online supplementary materials. Whenever
children living in the same household were tested, we tried to test
children in age order, from oldest to youngest, to minimize intimidation
for younger household members, and always using different elicitation
sets.

We fitted the child with a headset microphone (Shure SM10A or WH20 XLR
with a dynamic microphone on a headband, most children using the former)
that fed into the left channel of a Tascam DR40x digital audio recorder.
The headsets were designed for adult use and could not be comfortably
seated on many children's heads without a more involved adjustment
period. To minimize adjustment time, which was uncomfortable for some
children given the proximity of the experimenter and equipment, we
placed the headband on children's shoulders in these cases, carefully
adjusting the microphone's placement so that it was still close to the
child's mouth. A research assistant who spoke Yélî Dnye natively sat
next to the child throughout the task to provide instructions and, if
needed, encouragement. The research assistant coached the child
throughout the task to make sure that they understood what they were
expected to do. An experimenter (the first author) delivered the
pre-recorded stimuli to the research assistant and the child over
headphones.

The first phase of the experiment involved making sure the child
understood the task. We explained the task and then orally presented the
first practice item. At this point, many children did not say anything
in response, which triggered the following procedure: First, the
assistant insisted the child make a response. If the child still did not
say anything, the assistant said a real word and then asked the child to
repeat it, then another and another. If the child could repeat real
words correctly, we provided the first training item over headphones
again for children to repeat. Most children successfully started
repeating the items at this point, but a few needed further help. In
this case, the assistant modeled the behavior (i.e., the child and
assistant would hear the item again, and the assistant would repeat it;
then we would play the item again and ask the child to repeat it). A
small minority of children still failed to repeat the item at this
point. If so, we tried again with the second training item, at which
point some children demonstrated task understanding and could continue.
A fraction of the remaining children, however, failed to repeat this
second training item, as well as the third one, in which case we stopped
testing altogether (see Participants section for exclusions).

The second phase of the experiment involved running the child through
the list of test items randomly assigned to them. This was done in the
same manner as the practice items: the stimulus was played over the
headphones, and then the child repeated it aloud. NWR studies vary in
whether children are allowed to hear and/or repeat the item more than
one time. We had a fixed procedure for the test items (i.e., the
non-practice items) in which the child was alowed to make a further
attempts if their first attempt was judged erroneous in some way by the
assistant. The procedure worked as follows: When the child made an
attempt, the assistant indicated to the experimenter whether the child's
production was correct or not. If correct, the experimenter would
whisper this note of correct repetition into a separate headset that fed
into the right channel of the same Tascam recorder and the group moved
on to the next item. If not, the child was allowed to try again, with up
to five attempts allowed before moving on to the next item. Children
were not asked to make repetitions if they did not produce a first
attempt. In total, test sessions took approximately XX minutes on
average, with XX of those minutes attributed to practice and XX to the
actual test list.

@@

Additionally, we estimated the typological frequency of all phonological
segments used in the target items using the PHOIBLE cross-linguistic
phonological inventory database (Moran \& McCloy, 2019). For each phone
in our task, we extracted the number and percentage of languages noted
to have that phone in its inventory. While PHOIBLE is an unprecedentedly
comprehensive database, with phonological inventory data for over 2000
languages at time of writing, it is of course still far from complete,
so estimated frequencies should be taken with a pinch of salt. In both
PHOIBLE and UPSID---another large phonemic inventory database---note
that nearly half of the segment types are only attested in one language
(Steven Moran, personal communication; see also
\href{http://web.phonetik.uni-frankfurt.de/upsid_info.html}{web.phonetik.uni-frankfurt.de/upsid\_info.html}).
It is therefore unsurprising that we did not find three of our segments
(/lβʲ/, /ṭp/, and /tp/) in the database. Following the general pattern
that many phones only appear once cross-linguistically, even in large
databases, we treat our ``missing'' phones as having a frequency of 1,
with a (rounded) percentage of 0\%.

@@

We tested a total of 55 children from 38 families spread across four
hamlets. We excluded test sessions from analysis for the following
reasons: refused participation or failure to repeat items presented over
headphones even after coaching (N=8), spoke too softly to allow offline
coding (N=5), or were 13 years old or older (N=2; we tested these
teenagers to put younger children at ease). The remaining 40 children
(14 girls) were aged 6.98 years (range 3.92-11.03 years). Among these
children were 34 only exposed to Yélî Dnye at home and 6 children
exposed to Yélî Dnye plus one or more other languages at home. Maternal
years of education averaged 8.22 years (range 6-12 years).\footnote{Years
  of education is difficult to precisely pinpoint in this context.
  Schooling beyond grade 8 takes place off-island and enrollment is tied
  to an ability to pay that year's school fees, seasonal interruptions
  to school opening times, familial obligations, and more. We asked
  parents to report their highest completed level of education. We then
  record the number of years entailed by having completed that level
  under ideal conditions. This method results in many esimates around
  typical ``graduation'' times (e.g., graduating from primary education
  before starting secondary education).} In terms of birth order, 6 were
first borns, 5 second, 2 third, 7 forth, 5 fifth, and 1 sixth.

@@

We first checked whether whole-item NWR scores varied between first and
subsequent presentations of an item by averaging word-level scores at
the participant level separately for first attempts and subsequent
repetitions. We excluded 1 child who did not have data for one of these
two types. As shown in Figure \ref{fig:Fig1-first-vs-others},
participants' mean word-level scores became more heterogeneous in
subsequent repetitions. Surprisingly, whole-item NWR scores for
subsequent repetitions (M = 40, SD = 26) were on average lower than
first ones (M = 64, SD = 15), t(38) = 6.28, p \textless{} 0.001; Cohen's
d = 1.14). Given uncertainty in whether previous work used first or all
repetitions, and given that score here declined and became more
heterogeneous in subsequent repetitions, we focus the remainder of our
analyses only on first repetitions.

Taking into account only the first attempts, we averaged attempts by
each of the 40 children who had data for first attempts; their ages
ranged from 3 to 10 (M = 6.50, SD = 1.50). The overall NWR score was M =
64\% (SD = 15\%), Cohen's d = 4.36. Scores based on phonemes are even
higher M = 78\% (SD = 9\%), Cohen's d = 8.55. The phoneme-based
normalized Levenshtein distance was M = 22\% (SD = 9\%), meaning that
about a fifth of phonemes were were substituted, inserted, or deleted
(the normalized Levenshtein distance is the complement of phoneme-based
scores).

@@

We then analyzed variation in whole-item NWR scores as a function of the
average frequency with which sounds composing individual target words
are found in languages over the world. To look at this, we fit a mixed
logistic regression in which the outcome variable was whether the
non-word was correctly repeated or not. The fixed effect of interest was
the average cross-linguistic phone frequency; we also included child age
as a control fixed effect, and child ID \& target ID as random
variables.

We could include 850 observations, from 40 children producing in any
given trial one of 41 potential target words. The analysis revealed a
main effect of age (ß = 0.33, SE ß = 0.13, p = 0.01); and a significant
estimate for the scaled average cross-linguistic frequency of phones in
the target words (ß = 0.82, SE ß = 0.18, p \textless{} 0.001): Target
words with phones found more frequently across languages had a higher
proportion of words that were correctly repeated, as shown in Figure
\ref{fig:Fig2-xling-freq}. Averaging across items and participants, the
Pearson correlation between scaled average cross-linguistic phone
frequency and whole-item NWR scores was r(31) = 0.63.

@@

We next checked whether the association between whole-item NWR scores
and cross-linguistic phone frequence could actually be due to frequency
of the sounds within the language: One can suppose that sounds that
occur more frequently across languages are also more frequent within a
language, and therefore may be easier for children to represent and
repeat. We estimated frequency of the phones present in the stimuli in a
corpus of child-centered recordings (Casillas et al., 2020) by counting
the number of word types in which they occurred, and applied the natural
logarithm.\footnote{We also carried out this analysis using token phone
  frequency, but this measure was not correlated with whole-item NWR
  scores, and therefore the fact that it did not explain away the
  predictive value of cross-linguistic phone frequency was less
  informative than the relationship discussed in the main text, with
  type frequencies.} As before, unattested sounds were not considered
(i.e., they were declared NA so that they do not count for analyses).
Phone frequencies estimated from this corpus were correlated with
cross-linguistic phone frequencies (r(27)=0.52, p \textless{} 0.01).
Moreover, averaging across items and participants, the Pearson
correlation between scaled average corpus phone frequency and whole-item
NWR scores was r = 0.46. Therefore, we fit another mixed logistic
regression, this time declaring as fixed effects both scaled
cross-linguistic and corpus frequencies (averaged across all attested
phones within each stimulus item), in addition to age. As before, the
model contained random slopes for both child ID and target. In this
model, both cross-linguistic frequency (ß = 0.82, SE ß = 0.27, p
\textless{} 0.01) and age (ß = 0.33, SE ß = 0.13, p = 0.01) were
significant predictors of whole-item NWR scores, but corpus frequency (ß
= 0.00, SE ß = 0.25, p = 1) was not.

@@

Next, we investigated patterns of deletion and substitution. Deletions
were relatively rare: there were only 7 instances of deleted vowels
(\textasciitilde{}0.23\% of all vowel targets), and 3 instances of
deleted consonants (\textasciitilde{}0.12\% of all consonant targets).
Substitutions were somewhat more frequent, but still relatively rare.
Nasal vowels were produced as oral vowels and vice versa (34 oral
targets as nasal (\textasciitilde{}0.99\% of targets) and 30 nasal
targets produced as oral (\textasciitilde{}2.37\% of targets). Both oral
and nasal vowels were also occasionally subsituted by vowels of a
different quality that preserved the target oral/nature airflow (110
oral vowels (\textasciitilde{}1.86\% of targets and 13 nasal vowels
(\textasciitilde{}2.06\% of targets). Among consonants, complex
consonants were occasionally substituted by simpler ones (specifically
{[}ṭp{]}, {[}tp{]}, {[}kp{]}, {[}km{]}, {[}kṇ{]}, {[}mp{]}, {[}ɣ{]}, and
{[}lβʲ{]} as {[}m{]}, {[}n{]}, {[}l{]}, {[}w{]}, {[}j{]}, {[}w{]},
{[}ṭ{]}, {[}g{]}, {[}p{]}, {[}t{]}, {[}k{]}, {[}f{]}, {[}h{]}, and
{[}tʃ{]}), 115 times, in approximately 1.96\% of all complex consonant
targets. Interestingly, simple consonants were sometimes also
substituted by complex ones; this happened 2 times, in approximately
1.96\% of all simple consonant targets. Finally, simple consonants were
mispronounced as other simple consonants (68 instances,
\textasciitilde{}0.38\% of all simple consonant targets) and complex
consonants as other complex consonants (33 instances,
\textasciitilde{}1.02\% of all complex consonant targets). Note that the
substitutions included phones that are not native to Yélî Dnye but do
occur in English (e.g., {[}tʃ{]}); as these data come from careful
transcriptions by a native Yélî Dnye speaker who is very fluent in
English, we take these segmental substitutions faithfully as an
indication that several of our participants have mastered production of
some English phones, possibly produced within whole English word forms
(see below).

Finally, we looked into the frequency with which mispronunciations
resulted in real words. In fact, two thirds of incorrect repetitions
were recognizable as real words or phrases in Yélî Dnye or English:
63\%. This type of analysis is seldom reported---we could only find one
comparison point: ({\textbf{???}}) found that illiterate Portuguese
adults' NWR mispronunciations resulted in real words in 11.16\% of
cases, whereas literate participants did so in only 1.71\% of cases. The
percentage we observe here is much higher than reported in Castro and
colleagues' study, but we do not know whether age, language, or even
test structure explains this difference.

@@

Next, we inspected whether NWR scores varied as a function of word
length (Table \ref{tab:tablength}). Participants scored much lower on
monosyllables than on non-words of other lengths. This is likely due to
the fact that the majority of monosyllables were designed to include
sounds that are rare in the world's languages, which may indicate that
they are hard to produce or perceive. Setting monosyllables aside, we
observe the typical pattern of lower scores for longer items, although
this is particularly salient for the whole-item scoring. While
whole-item scoring is the most commonly reported NWR outcome, it is also
the least forgiving. The pattern is less marked when other two scores
are used, which are less sensitive to errors. Averaging across
participants and items, the Pearson correlation between length (2-4
syllables) and whole-item NWR scores was r(1) = -0.91. In a generalized
binomial mixed model, we included 479 observations, from 40 children
producing, in any given trial, one of 24 potential target words. The
analysis revealed a positive effect of age (ß = 0.56, SE ß = 0.14, p
\textless{} 0.001) and a negative but non-significant estimate for
target length in number of syllables (ß = -0.15, SE ß = 0.33, p = 0.65).

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Fig2-scores-by-age-1.pdf}
\caption{\label{fig:Fig2-scores-by-age}NWR whole-item scores for individual
participants as a function of age and sex (purple = boys, orange =
girls).}
\end{figure}

Our final exploratory analysis assessed whether variation in scores was
structured by factors that vary across individuals. As shown in Figure
\ref{fig:Fig2-scores-by-age}, there was a greater deal of variance
across the tested age range, with significantly higher NWR scores for
older children (Spearman's rank correlation, given inequality of
variance, rho (6,014.70) = 0.44, p \textless{} 0.01). In contrast, there
was no clear association between NWR scores and sex (t (-0.29) = 27.56,
p = 0.77), birth order (data missing for 15 children, rho = (3,441.90) =
-0.18, p = 0.39), or maternal education (data missing for 0 children,
rho ( 9,594.37) = 0.10, p = 0.54).
